Download data

### 2023-12-29 scripts/download_independent_variables.Rmd
adding empty scripts for steps
outlier detection methods
stuff

```{r}
library(tidyverse)
```

The main bioclim variables can be found at the [link here](https://envicloud.wsl.ch/#/?bucket=https%3A%2F%2Fos.zhdk.cloud.switch.ch%2Fenvicloud%2F&prefix=chelsa%2Fchelsa_V2%2FGLOBAL%2Fclimatologies%2F1981-2010%2Fbio%2F) an S3 bucket for CHELS*A* (not the neighborhood!).
All files can be downloaded as a textfile for further use with wget. 
We will subset these files to only download the layers of interest.

@@

#### Soil USDA Class ####

@@

>>>>

```{r}
library(tidyverse)
library(terra)
```

The main bioclim variables can be found at the [link here](https://envicloud.wsl.ch/#/?bucket=https%3A%2F%2Fos.zhdk.cloud.switch.ch%2Fenvicloud%2F&prefix=chelsa%2Fchelsa_V2%2FGLOBAL%2Fclimatologies%2F1981-2010%2Fbio%2F) an S3 bucket for CHELS*A* (not the neighborhood!).
All files can be downloaded as a textile for further use with wget. 
We will subset these files to only download the layers of interest.

@@

@@

process dem
```{r}
p2dem <- '/media/steppe/hdd/Geospatial_data/merit-dem'
files <- file.path(p2dem, list.files(p2dem, recursive = TRUE, pattern= 'flt'))
files <- files[15:length(file)] # lower files south of extent

dem <- vrt(files)
dem <- crop(dem, project(domain, crs(dem)))
dem <- project(dem, 'EPSG:5070')

pout <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/dem/' 
filename <- paste0(pout, 'dem_.tif')

bio1 <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/chelsa/bio1'
target_cells <- vrt(file.path(bio1, list.files(bio1)))
dem <- resample(dem, target_cells)

template <- rast(nrows = 25, ncols = 25) # create a big empty raster, you can go in through sf too. 
ext(template) <- c( -125.5, -100, 27,  50) # set the extent
crs(template) <- "EPSG:4326" 
template <- project(template, 'EPSG:5070')

makeTiles(dem, template, filename)

rm(p2dem, files, dem, pout, filename, bio1, target_cells, template)
```
